Calendar-based graphics for visualizing people's daily schedules

### 2019-10-13 reviews/response.Rmd

> Page 19, first sentence. The meaning of this is not clear. Rephrase.

Done. We rephrased to "The previous calendar plots are static, made with **ggplot2**."

> Page 21, figure 12. I agree with reviewer 1 that this needs some tweaking. Too much vertical space given to panel labels. Time labels are confusing. Two of the line colours are very similar here and in other plots.

We've reduced the size of panel labels to give more room to the data display. Time labels are now made to show 6 and 18 to avoid confusion.

The colour scheme is suggested by reviewer 3 before, which is colour-blind and grey-scale print friendly.

@@

> The authors mention "modulus arithmetic" three times and linear algebra once, so I expected complicated mathematics.  The mathematics is simple.  These remarks are not needed.  Most people say "modular arithmetic".

Done.

> p6 "Jones (2016), Wong (2013), Kothari & Ather (2016), and Jacobs (2017) implemented some variants of calendar-based heatmaps as in R packages: TimeProjection, ggTimeSeries, and ggcal respectively."  Four articles, three packages, one must be missing.  Better when references are next to package names.

We removed the Jones (2016), since it's an online post. We also rephrased the sentence so that references are next to package names.

@@

> p7 "Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default."  This is not true for six-week months.

Thanks. We revised this sentence to "Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to the week of month and the day of week respectively."

@@

> p19 Text implies that more than one date can be selected in the Shiny app.  The app was kludgy and did not work smoothly.  It is a nice idea, but interaction has to be fast and work well.  Zooming did not work properly and was not able to link from the weather plots.  Fix or do not mention in article.

TODO

> p21 Figure 12 is an interesting possibility, but not well implemented. The repetition of the scales in unhelpful. 0 12 24 looks like 240 12 240.  There is too much labelling and too little information.

We've reduced the size of panel labels to give more room to the data display. Time labels are now made to show 6 and 18 to avoid confusion.

@@

> The case study is interesting.  Why are the data for household 2 different (fewer possible values)?

Data is of the same time lengths (and the same amount of data points) across four households. Different households have different lifestyles.

> p22 "with noticeable variations on Thursdays" based on only 26 data points is over-interpretation.

### 2019-10-13 calendar-vis.Rmd
tweak figure 12

In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events. There are some existing works in this area. For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days. It contrasts weekdays and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office. @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively. However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day. Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation. Color-encoding is also low on the hierarchy of optimal variable mapping [@cleveland1984graphical; @lam2007overview].

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where the top left is Sunday of week 1, and the bottom right is Monday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The notation for creating these cells is as follows:

>>>>

In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events. There are some existing works in this area. For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days. It contrasts weekdays and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office. Some variants of calendar-based heatmaps have been implemented in R packages: **TimeProjection** [@R-TimeProjection], **ggTimeSeries** [@R-ggTimeSeries], and **ggcal** [@R-ggcal]. However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day. Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation. Color-encoding is also low on the hierarchy of optimal variable mapping [@cleveland1984graphical; @lam2007overview].

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to the week of month and the day of week respectively. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The notation for creating these cells is as follows:
